---
title: "model county exploration"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r}
library(dplyr)
library(ggplot2)
library(GGally)
library(gridExtra)
```


Assumption, Cameron, East Faleciana, Iberville, La Salle, Red
River, and West Baton Rouge counties.

Other features: health factors rank, life quality rank. health be-
havior rank, adult smoking %, food environment index, percentage of uninsured
people, rate of completion for post-secondary education, percentage of diabetic people

Predictor: adult_obesity_percent

Use the obesity data set

```{r}
imp_features <- c("County_name", "Year", "hispanic_percent", "african_american_percent", "nonhispanic_white_percent","rural","female","health_factors_rank", "life_quality_rank", "health_behav_rank", "adult_smoking_percent", "adult_obesity_percent","food_env_index", "uninsured_percent", "PSED_completion_rate", "diabetic_percent")

counties <- c("Assumption", "Cameron", "East Feliciana","Iberville",  "La Salle", "Red River", "West Baton Rouge")
```

Assumption, Cameron, East Faleciana, Iberville, La Salle, Red
River, and West Baton Rouge

create separate data sets for each county

```{r}
assumption <- obesity[imp_features] %>%
  filter(County_name == "Assumption")

cameron <- obesity[imp_features] %>%
  filter(County_name == "Cameron")

east_faleciana <- obesity[imp_features] %>%
  filter(County_name == "East Feliciana")

iberville <- obesity[imp_features] %>%
  filter(County_name == "Iberville")

la_salle <- obesity[imp_features] %>%
  filter(County_name == "La Salle")

red_river <- obesity[imp_features] %>%
  filter(County_name == "Red River")

west_b_rouge <- obesity[imp_features] %>%
  filter(County_name == "West Baton Rouge")

imp_counties <- obesity[imp_features] %>%
  filter(County_name %in% counties)

nrow(imp_counties)

```


how obesity and smoking interact in the 7 counties
```{r}
ggplot(imp_counties, aes(adult_smoking_percent, adult_obesity_percent, color = factor(Year))) + 
  geom_point() +
  geom_smooth(method = "loess")
```

```{r}
ggplot(imp_counties, aes(adult_smoking_percent, adult_obesity_percent, color = factor(County_name))) + 
  geom_point() +
  geom_smooth(method = "loess")
```

Other features: health factors rank, life quality rank. health be-
havior rank, adult smoking %, food environment index, percentage of uninsured
people, rate of completion for post-secondary education, percentage of diabetic people

Predictor: adult_obesity_percent

getting avergages for the non county variables, for each county
```{r}
counties <- group_by(imp_counties, County_name)
avg_smoking <- summarize(counties,avg= mean(adult_smoking_percent,na.rm=TRUE))
avg_obesity <- summarize(counties,avg= mean(adult_obesity_percent,na.rm=TRUE))
avg_food_index <- summarize(counties,avg= mean(food_env_index,na.rm=TRUE))
avg_uninsured <- summarize(counties,avg= mean(uninsured_percent,na.rm=TRUE))
avg_psed <- summarize(counties,avg= mean(PSED_completion_rate,na.rm=TRUE))
avg_diabetic <- summarize(counties,avg= mean(diabetic_percent,na.rm=TRUE))

avg_obesity
avg_smoking
avg_food_index
avg_uninsured
avg_psed
avg_diabetic
```


plots for all the non rank variables, including dependent variable
```{r}
ggplot(imp_counties, aes(factor(Year), adult_obesity_percent, color = factor(County_name))) + 
  geom_point() +
  geom_line(aes(group = factor(County_name)))+
  labs(title ="Change in adult obesity from 2010 to 2020", x = "Year", y = "Adult obesity %") 
  #geom_smooth(method = "loess")
```

```{r}
ggplot(imp_counties, aes(factor(Year), adult_smoking_percent, color = factor(County_name))) + 
  geom_point() +
  geom_line(aes(group = factor(County_name)))+
  labs(title ="Change % of uninsured people", x = "Year", y = "Uninsured %") 
  #geom_smooth(method = "loess")
```

The food environment index combines two measures of food access: the percentage of the population that is low-income and has low access to a grocery store, and the percentage of the population that did not have access to a reliable source of food during the past year (food insecurity).

```{r}
ggplot(imp_counties, aes(factor(Year), food_env_index, color = factor(County_name))) + 
  geom_point() +
  geom_line(aes(group = factor(County_name)))+
  labs(title ="Change % of uninsured people", x = "Year", y = "Uninsured %") 
  #geom_smooth(method = "loess")
```

```{r}
ggplot(imp_counties, aes(factor(Year), PSED_completion_rate, color = factor(County_name))) + 
  geom_point() +
  geom_line(aes(group = factor(County_name)))+
  labs(title ="Change in PSED completion rate", x = "Year", y = "PSED completion rate") 
  #geom_smooth(method = "loess")
```

```{r}
ggplot(imp_counties, aes(factor(Year), uninsured_percent, color = factor(County_name))) + 
  geom_point() +
  geom_line(aes(group = factor(County_name)))+
  labs(title ="Change % of uninsured people", x = "Year", y = "Uninsured %") 
  #geom_smooth(method = "loess")
```

```{r}
ggplot(imp_counties, aes(factor(Year), diabetic_percent, color = factor(County_name))) + 
  geom_point() +
  geom_line(aes(group = factor(County_name)))+
  labs(title ="Change in % of diabetic people", x = "Year", y = "% of diabetic people") 
  #geom_smooth(method = "loess")
```

health factors rank, life quality rank. health behavior rank

"health_factors_rank", "life_quality_rank", "health_behav_rank"

plots for ranked variables

```{r}
ggplot(imp_counties, aes(factor(Year), health_factors_rank, color = factor(County_name))) + 
  geom_point() +
  geom_line(aes(group = factor(County_name)))+
  labs(title ="Change in health factors ranks", x = "Year", y = "Health factors rank") 
  #geom_smooth(method = "loess")
```

```{r}
ggplot(imp_counties, aes(factor(Year), life_quality_rank, color = factor(County_name))) + 
  geom_point() +
  geom_line(aes(group = factor(County_name)))+
  labs(title ="Change in life quality ranks", x = "Year", y = "Life quality rank") 
  #geom_smooth(method = "loess")
```

```{r}
ggplot(imp_counties, aes(factor(Year), health_behav_rank, color = factor(County_name))) + 
  geom_point() +
  geom_line(aes(group = factor(County_name)))+
  labs(title ="Change in health behavior ranks", x = "Year", y = "Health behavior rank") 
  #geom_smooth(method = "loess")
```


plots for racial demographics

```{r}
ggplot(imp_counties, aes(factor(Year), african_american_percent, color = factor(County_name))) + 
  geom_point() +
  geom_line(aes(group = factor(County_name)))
  #geom_smooth(method = "loess")
```

```{r}
ggplot(imp_counties, aes(factor(Year), hispanic_percent, color = factor(County_name))) + 
  geom_point() +
  geom_line(aes(group = factor(County_name)))
  #geom_smooth(method = "loess")
```

```{r}
ggplot(imp_counties, aes(factor(Year), nonhispanic_white_percent, color = factor(County_name))) + 
  geom_point() +
  geom_line(aes(group = factor(County_name)))
  #geom_smooth(method = "loess")
```



*****************************Individual plots**********************************************************************

change in obesity over time
```{r}
a <- ggplot(red_river, aes(factor(Year), adult_obesity_percent, color = factor(County_name))) + 
  geom_point() +
   geom_line(aes(group = factor(County_name)))
  #geom_smooth(method = "loess", formula = "y~x")


b <- ggplot(cameron, aes(factor(Year), adult_obesity_percent, color = factor(County_name))) + 
  geom_point() +
   geom_line(aes(group = factor(County_name)))
  #geom_smooth(method = "loess", formula = "y~x")

c <- ggplot(assumption, aes(factor(Year), adult_obesity_percent, color = factor(County_name))) + 
  geom_point() +
  geom_line(aes(group = factor(County_name)))
  #geom_smooth(method = "loess", formula = "y~x")

d <- ggplot(east_faleciana, aes(factor(Year), adult_obesity_percent, color = factor(County_name))) + 
  geom_point() +
   geom_line(aes(group = factor(County_name)))
  #geom_smooth(method = "loess", formula = "y~x")

e <- ggplot(iberville, aes(factor(Year), adult_obesity_percent, color = factor(County_name))) + 
  geom_point() +
  geom_line(aes(group = factor(County_name)))
  #geom_smooth(method = "loess", formula = "y~x")

f <- ggplot(la_salle, aes(factor(Year), adult_obesity_percent, color = factor(County_name))) + 
  geom_point() +
  geom_line(aes(group = factor(County_name)))
  #geom_smooth(method = "loess", formula = "y~x")

g <- ggplot(west_b_rouge, aes(factor(Year), adult_obesity_percent, color = factor(County_name))) + 
  geom_point() +
  geom_line(aes(group = factor(County_name)))
  #geom_smooth(method = "loess", formula = "y~x")

all_obesity <- grid.arrange(a, b, c, d, e, f, g, nrow = 4, ncol = 2)
all_obesity
```


change in smoking over time
```{r}
a1 <- ggplot(red_river, aes(factor(Year), adult_smoking_percent, color = factor(County_name))) + 
  geom_point() +
   geom_line(aes(group = factor(County_name)))
  #geom_smooth(method = "loess", formula = "y~x")


b1 <- ggplot(cameron, aes(factor(Year), adult_smoking_percent, color = factor(County_name))) + 
  geom_point() +
   geom_line(aes(group = factor(County_name)))
  #geom_smooth(method = "loess", formula = "y~x")

c1 <- ggplot(assumption, aes(factor(Year), adult_smoking_percent, color = factor(County_name))) + 
  geom_point() +
  geom_line(aes(group = factor(County_name)))
  #geom_smooth(method = "loess", formula = "y~x")

d1 <- ggplot(east_faleciana, aes(factor(Year), adult_smoking_percent, color = factor(County_name))) + 
  geom_point() +
   geom_line(aes(group = factor(County_name)))
  #geom_smooth(method = "loess", formula = "y~x")

e1 <- ggplot(iberville, aes(factor(Year), adult_smoking_percent, color = factor(County_name))) + 
  geom_point() +
  geom_line(aes(group = factor(County_name)))
  #geom_smooth(method = "loess", formula = "y~x")

f1 <- ggplot(la_salle, aes(factor(Year), adult_smoking_percent, color = factor(County_name))) + 
  geom_point() +
  geom_line(aes(group = factor(County_name)))
  #geom_smooth(method = "loess", formula = "y~x")

g1 <- ggplot(west_b_rouge, aes(factor(Year), adult_smoking_percent, color = factor(County_name))) + 
  geom_point() +
  geom_line(aes(group = factor(County_name)))
  #geom_smooth(method = "loess", formula = "y~x")

all_smoking <- grid.arrange(a1, b1, c1, d1, e1, f1, g1, nrow = 4, ncol = 2)
all_smoking
```

avg_food_index <- summarize(counties,avg= mean(food_env_index,na.rm=TRUE))
avg_uninsured <- summarize(counties,avg= mean(uninsured_percent,na.rm=TRUE))
avg_psed <- summarize(counties,avg= mean(PSED_completion_rate,na.rm=TRUE))
avg_diabetic <- summarize(counties,avg= mean(diabetic_percent,na.rm=TRUE))

change in food environment index over time
```{r}
a2 <- ggplot(red_river, aes(factor(Year), food_env_index, color = factor(County_name))) + 
  geom_point() +
   geom_line(aes(group = factor(County_name)))
  #geom_smooth(method = "loess", formula = "y~x")


b2 <- ggplot(cameron, aes(factor(Year), food_env_index, color = factor(County_name))) + 
  geom_point() +
   geom_line(aes(group = factor(County_name)))
  #geom_smooth(method = "loess", formula = "y~x")

c2 <- ggplot(assumption, aes(factor(Year), food_env_index, color = factor(County_name))) + 
  geom_point() +
  geom_line(aes(group = factor(County_name)))
  #geom_smooth(method = "loess", formula = "y~x")

d2 <- ggplot(east_faleciana, aes(factor(Year), food_env_index, color = factor(County_name))) + 
  geom_point() +
   geom_line(aes(group = factor(County_name)))
  #geom_smooth(method = "loess", formula = "y~x")

e2 <- ggplot(iberville, aes(factor(Year), food_env_index, color = factor(County_name))) + 
  geom_point() +
  geom_line(aes(group = factor(County_name)))
  #geom_smooth(method = "loess", formula = "y~x")

f2 <- ggplot(la_salle, aes(factor(Year), food_env_index, color = factor(County_name))) + 
  geom_point() +
  geom_line(aes(group = factor(County_name)))
  #geom_smooth(method = "loess", formula = "y~x")

g2 <- ggplot(west_b_rouge, aes(factor(Year), food_env_index, color = factor(County_name))) + 
  geom_point() +
  geom_line(aes(group = factor(County_name)))
  #geom_smooth(method = "loess", formula = "y~x")

all_foodenvindex <- grid.arrange(a2, b2, c2, d2, e2, f2, g2, nrow = 4, ncol = 2)
all_foodenvindex
```

change in uninsured people over time
```{r}
a3 <- ggplot(red_river, aes(factor(Year), uninsured_percent, color = factor(County_name))) + 
  geom_point() +
   geom_line(aes(group = factor(County_name)))
  #geom_smooth(method = "loess", formula = "y~x")


b3 <- ggplot(cameron, aes(factor(Year), uninsured_percent, color = factor(County_name))) + 
  geom_point() +
   geom_line(aes(group = factor(County_name)))
  #geom_smooth(method = "loess", formula = "y~x")

c3 <- ggplot(assumption, aes(factor(Year), uninsured_percent, color = factor(County_name))) + 
  geom_point() +
  geom_line(aes(group = factor(County_name)))
  #geom_smooth(method = "loess", formula = "y~x")

d3 <- ggplot(east_faleciana, aes(factor(Year), uninsured_percent, color = factor(County_name))) + 
  geom_point() +
   geom_line(aes(group = factor(County_name)))
  #geom_smooth(method = "loess", formula = "y~x")

e3 <- ggplot(iberville, aes(factor(Year), uninsured_percent, color = factor(County_name))) + 
  geom_point() +
  geom_line(aes(group = factor(County_name)))
  #geom_smooth(method = "loess", formula = "y~x")

f3 <- ggplot(la_salle, aes(factor(Year), uninsured_percent, color = factor(County_name))) + 
  geom_point() +
  geom_line(aes(group = factor(County_name)))
  #geom_smooth(method = "loess", formula = "y~x")

g3 <- ggplot(west_b_rouge, aes(factor(Year), uninsured_percent, color = factor(County_name))) + 
  geom_point() +
  geom_line(aes(group = factor(County_name)))
  #geom_smooth(method = "loess", formula = "y~x")

all_uninsured <- grid.arrange(a3, b3, c3, d3, e3, f3, g3, nrow = 4, ncol = 2)
all_uninsured
```

change in PSED completion over time
```{r}
a4 <- ggplot(red_river, aes(factor(Year), PSED_completion_rate, color = factor(County_name))) + 
  geom_point() +
   geom_line(aes(group = factor(County_name)))
  #geom_smooth(method = "loess", formula = "y~x")


b4 <- ggplot(cameron, aes(factor(Year), PSED_completion_rate, color = factor(County_name))) + 
  geom_point() +
   geom_line(aes(group = factor(County_name)))
  #geom_smooth(method = "loess", formula = "y~x")

c4 <- ggplot(assumption, aes(factor(Year), PSED_completion_rate, color = factor(County_name))) + 
  geom_point() +
  geom_line(aes(group = factor(County_name)))
  #geom_smooth(method = "loess", formula = "y~x")

d4 <- ggplot(east_faleciana, aes(factor(Year), PSED_completion_rate, color = factor(County_name))) + 
  geom_point() +
   geom_line(aes(group = factor(County_name)))
  #geom_smooth(method = "loess", formula = "y~x")

e4 <- ggplot(iberville, aes(factor(Year), PSED_completion_rate, color = factor(County_name))) + 
  geom_point() +
  geom_line(aes(group = factor(County_name)))
  #geom_smooth(method = "loess", formula = "y~x")

f4 <- ggplot(la_salle, aes(factor(Year), PSED_completion_rate, color = factor(County_name))) + 
  geom_point() +
  geom_line(aes(group = factor(County_name)))
  #geom_smooth(method = "loess", formula = "y~x")

g4 <- ggplot(west_b_rouge, aes(factor(Year), PSED_completion_rate, color = factor(County_name))) + 
  geom_point() +
  geom_line(aes(group = factor(County_name)))
  #geom_smooth(method = "loess", formula = "y~x")

all_psed <- grid.arrange(a4, b4, c4, d4, e4, f4, g4, nrow = 4, ncol = 2)
all_psed
```

change in diabetes prevalence over time
```{r}
a5 <- ggplot(red_river, aes(factor(Year), diabetic_percent, color = factor(County_name))) + 
  geom_point() +
   geom_line(aes(group = factor(County_name)))
  #geom_smooth(method = "loess", formula = "y~x")


b5 <- ggplot(cameron, aes(factor(Year), diabetic_percent, color = factor(County_name))) + 
  geom_point() +
   geom_line(aes(group = factor(County_name)))
  #geom_smooth(method = "loess", formula = "y~x")

c5 <- ggplot(assumption, aes(factor(Year), diabetic_percent, color = factor(County_name))) + 
  geom_point() +
  geom_line(aes(group = factor(County_name)))
  #geom_smooth(method = "loess", formula = "y~x")

d5 <- ggplot(east_faleciana, aes(factor(Year), diabetic_percent, color = factor(County_name))) + 
  geom_point() +
   geom_line(aes(group = factor(County_name)))
  #geom_smooth(method = "loess", formula = "y~x")

e5 <- ggplot(iberville, aes(factor(Year), diabetic_percent, color = factor(County_name))) + 
  geom_point() +
  geom_line(aes(group = factor(County_name)))
  #geom_smooth(method = "loess", formula = "y~x")

f5 <- ggplot(la_salle, aes(factor(Year), diabetic_percent, color = factor(County_name))) + 
  geom_point() +
  geom_line(aes(group = factor(County_name)))
  #geom_smooth(method = "loess", formula = "y~x")

g5 <- ggplot(west_b_rouge, aes(factor(Year), diabetic_percent, color = factor(County_name))) + 
  geom_point() +
  geom_line(aes(group = factor(County_name)))
  #geom_smooth(method = "loess", formula = "y~x")

all_diabetic <- grid.arrange(a5, b5, c5, d5, e5, f5, g5, nrow = 4, ncol = 2)
all_diabetic
```



