---
title: "data pre-processing"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


#packages required

```{r}
#install.packages("gridExtra")
library(readr)
library(dplyr)
library(tidyr)
library(ggplot2)
library(gridExtra)
#install.packages("GGally")
library(GGally)
#install.packages("caret")
library(caret)

#install.packages("Rtools")
#library(Rtools)
```

read in data
```{r}
data_obesity <- read_csv("data_obesity.csv")
```

#remove columns below
```{r}
data_obesity <- subset(data_obesity, select = -c(X68, X69, X70, X71, X72))
data_obesity <- subset(data_obesity, select = -c(Year_1))

```

#remove rows where county is na
```{r}
obesity <- data_obesity %>%
  drop_na(County)
```


```{r}
str(obesity)
```

```{r}
summary(obesity)
```



dealing with missing data
```{r}
sum(is.na(obesity)) #14570

#gives sum of na for all columns 
obesity %>%
  summarise_all(funs(sum(is.na(.))))

#thoughts: most of the variables that will go in the model won't have a lot of NAs. This is because all the ones chosen for the big model are chosen based on when they were introduced. Can't do anything about those NAs tbh because replacing them with something else won't reflect the reality of the data at that time period. 

```


dealing with categorical data: county

```{r}
#dupicating the county column
obesity %>% 
  mutate(County_name = County)

#change county column to a factor
obesity$County <- as.numeric(factor(obesity$County))
```


looking for outliers-- univariate approach
```{r}
ggplot(obesity, aes(x = adult_obesity_percent))+ 
  geom_histogram() +
  stat_bin(bins = 20, binwidth = 1) +
  geom_vline(aes(xintercept=mean(adult_obesity_percent)),
            color="blue", linetype="dashed", size=1)
```

```{r}
boxplot(obesity$adult_obesity_percent,
  ylab = "adult_obesity_percent"
)
```


extract value of outliers based on the IQR criterion
```{r}
boxplot.stats(obesity$adult_obesity_percent)$out
```

use which() function to extract the row number corresponding to these outliers
```{r}
out <- boxplot.stats(obesity$adult_obesity_percent)$out
out_ind <- which(obesity$adult_obesity_percent %in% c(out))
out_ind

```

now easily go back to the specific rows in the dataset to verify them, or print all variables for these outliers:
```{r}
obesity[out_ind, ]
```

formal techniques to test for outliers: 2 statistical tests- grubb's (only one outlier at a time) and rosner (good for data sets with n> 25 and more than one outlier)

Grubb's test: allows to detect whether the highest or lowest value in a dataset is an outlier
```{r}
library(outliers)

# for lowest value
grubb_test <- grubbs.test(obesity$adult_obesity_percent)
grubb_test
```

```{r}
# for highest value
grubb_test <- grubbs.test(obesity$adult_obesity_percent, opposite = TRUE)
grubb_test
```

Rosner's test
```{r}
library(EnvStats)

rosner_test <- rosnerTest(obesity$adult_obesity_percent,
  k = 5 #no. of suspected outliers
)
rosner_test
```



multivariate outlier detection approach-- using cook's distance
```{r}
vars <- c("County", "rural", "african_american_percent", "american_indian_percent", "asian_percent", "pacific_islander_percent", "hispanic_percent", "nonhispanic_white_percent", "not_eng_prof_percent", "female", "health_outcomes_rank", "health_factors_rank", "life_length_rank", "life_quality_rank", "health_behav_rank", "clinical_care_rank", "socio_econ_fact_rank", "physical_env_rank", "premature_death_ypll_rate ", "poor_or_fair_health_percent", "poor_physical_health_days_avg", "adult_smoking_percent", "excessive_drinking_percent", "adult_obesity_percent", "food_env_index", "diabetic_percent", "percent_receiving_hba1c", " percent_limited_access_healthy_food_access", "percent_food_insecure", "access_exercise_opp_percent", "physical_inactivity_percent", "hs_grad_rate", "PSED_completion_rate", "uninsured_percent", "unemployed_percent", "income_ineq_income_ratio", "median_household_income")

obs_mod <- obesity[, (colnames(obesity) %in% vars)]
```

```{r}
mod <- lm(adult_obesity_percent ~ ., data= obs_mod)
cooksd <- cooks.distance(mod)
```

```{r}
plot(cooksd, pch="*", cex=2, main="Influential Obs by Cooks distance")  # plot cook's distance
abline(h = 4*mean(cooksd, na.rm=T), col="red")  # add cutoff line
text(x=1:length(cooksd)+1, y=cooksd, labels=ifelse(cooksd>4*mean(cooksd, na.rm=T),names(cooksd),""), col="red")  # add labels
```

```{r}
influential <- as.numeric(names(cooksd)[(cooksd > 4*mean(cooksd, na.rm=T))])  # influential row numbers
head(obs_mod[influential, ])  # influential observations.
```

Decision is to not remove any suspected outliers because each row is the value for a county and we expect there to be large variation between county ranks, demographics etc. 



normalize/standardize/scale the features
```{r}
#some plots before normalization etc
smoking <- ggplot(obesity, aes(adult_smoking_percent,adult_obesity_percent)) +
              geom_point()

alcohol <- ggplot(obesity, aes(excessive_drinking_percent,adult_obesity_percent)) +
              geom_point()

uninsured <- ggplot(obesity, aes(uninsured_percent,adult_obesity_percent)) +
              geom_point()

diabetes <- ggplot(obesity, aes(diabetic_percent, adult_obesity_percent)) +
              geom_point()

ypll <- ggplot(obesity, aes(premature_death_ypll_rate,adult_obesity_percent)) +
              geom_point()

income <- ggplot(obesity, aes(median_household_income,adult_obesity_percent)) +
              geom_point()

food_env <- ggplot(obesity, aes(food_env_index,adult_obesity_percent)) +
              geom_point()

grid.arrange(smoking, alcohol, uninsured, diabetes, ypll, income, food_env, nrow = 3, ncol = 3)
```

```{r}
#duplicate data set
obesity_norm <- data.frame(obesity)

#check it occupies a different memory location and hence doesn't point to the original data frame.
tracemem(obesity_norm)==tracemem(obesity)
```

standardization function
```{r}
standardize <- function(x){
  return((x - mean(x, na.rm = TRUE))/(sd(x, na.rm = TRUE)))
}
```

```{r}
#to get index of each column
data.frame(colnames(obs_mod))
```

stnadardize non rank columns
```{r}
obesity_standardize <- as.data.frame(apply(obs_mod[10:35],2,standardize)) #has all the non rank columns

```

standardize rank columns
```{r}
rank_standardize <- function(x){
  main = (x - min(x))/(max(x) - min(x))
  return((9*main) + 1)
}

#source: https://datascience.stackexchange.com/questions/1240/methods-for-standardizing-normalizing-different-rank-scales
```

```{r}
obesity_rank_standardize <- as.data.frame(apply(obs_mod[2:9],2,rank_standardize))
```

check before adding county
```{r}
dim(obesity_standardize)#704 26
dim(obesity_rank_standardize)#704 8
```


join all the standardized data: rank standardized, non rank standardized, and the county variable
```{r}
#add county to both standardized datasets:
obesity_standardize <- cbind(obesity_standardize, County = obs_mod$County)
obesity_rank_standardize <- cbind(obesity_rank_standardize, County = obs_mod$County)

#to confirm that county was added
dim(obesity_standardize)#704 27
dim(obesity_rank_standardize)#704 9

#View(obesity_rank_standardize)
#View(obesity_standardize)
```

cannot figure this out for the life of me:
```{r}
#full outer join on both the standardized datasets
obesity_all_stan <- merge(x = obesity_rank_standardize, y = obesity_standardize, by="County")#, all=TRUE)

#obesity_all_stan <- obesity_rank_standardize %>% 
 # inner_join(obesity_standardize, by="County")

#obesity_all_stan <- obesity_all_stan  %>% 
 # distinct(health_outcomes_rank, .keep_all = TRUE)

dim(obesity_all_stan) #704 35

```

for some reason I have 11 times the rows I did first. Each county hould have 11 entries and as there are 64 counties I should have 64*11 = 704 rows. But here there are 704*11 = 7744 rows. 
Each county appears a 121 times when it should appear 11 times. so it appears 11 times more than it should



New plan: export the standardized datasets from R and join them in excel
```{r}
write.csv(obesity_rank_standardize,'obesity_rank_standardize.csv')
write.csv(obesity_standardize,'obesity_standardize.csv')
```

```{r}
library(readxl)
obs_stan_all_use <- read_excel("obs_stan_all_use.xlsx")
#View(obs_stan_all_use)

#NA seen as character/string so wanted to make it this
obs_stan_all_use[obs_stan_all_use == "NA"] <- NA

#still the variables with NA are seen as characters
str(obs_stan_all_use)

#make it a dataframe
obs_stan_all_use <- as.data.frame(obs_stan_all_use)
```

```{r}
#made all the columns in the dataframe numeric
obs_stan_all_use <− lapply(obs_stan_all_use, as.numeric)
str(obs_stan_all_use)
```
make county a factor
```{r}
#county is a factor and should be treated as such
obs_stan_all_use$County <- as.factor(obs_stan_all_use$County)

str(obs_stan_all_use)
```

make dummy variables of county
```{r}
#install.packages("fastDummies")
library(fastDummies)

obs_stan_all_use <- dummy_cols(obs_stan_all_use, select_columns = 'County')

dim(obs_stan_all_use) #35 originally, 64 added, one for each county
```

```{r}
#remove the County variable
obs_stan_all_use <- subset(obs_stan_all_use, select= -c(County))

dim(obs_stan_all_use)#to confirm if column was removed
```


feature selection

```{r}
#for index
data.frame(colnames(obs_stan_all_use))
```

```{r}
#remove redundant features: remove attributes with an absolute correlation of 0.75 or higher

set.seed(5968)

install.packages("mlbench")
library(mlbench)
library(caret)

correlationMatrix <- cor(obs_stan_all_use[,1:34])
# summarize the correlation matrix
print(correlationMatrix)
# find attributes that are highly corrected (ideally >0.75)
highlyCorrelated <- findCorrelation(correlationMatrix, cutoff=0.75)
# print indexes of highly correlated attributes
print(highlyCorrelated)
```

so index 2, 1, 7, 4 ,5 are highly correlated at cutoff 0.5
2, 1 are highly correlated at 0.75

remove health_outcomes_rank
```{r}
obs_stan_all_use <- subset(obs_stan_all_use, select= -c(health_outcomes_rank))

dim(obs_stan_all_use)#to confirm if column was removed
```

```{r}
correlationMatrix <- cor(obs_stan_all_use[,1:97])
# summarize the correlation matrix
print(correlationMatrix)
# find attributes that are highly corrected (ideally >0.75)
highlyCorrelated <- findCorrelation(correlationMatrix, cutoff=0.75)
# print indexes of highly correlated attributes
print(highlyCorrelated)
```
1, 20 and 25 are highly correlated at 0.75. That's health_factors_rank, percent_food_insecure, african_american_percent. let it stay for now

```{r}
sum(is.na(obs_stan_all_use))
```

```{r}
summary(obs_stan_all_use)
```

replace all NAs with 0-- b/c it doesn't mean anything in the scaled data
```{r}
obs_stan_all_use[is.na(obs_stan_all_use)] <- 0
sum(is.na(obs_stan_all_use))#to confirm if all NAs were replaced
```


note: my current method disregards the year. so right now I will do it by 80/20 and later think about the years

splitting the data
```{r}
set.seed(78965)
#train <- obs_stan_all_use %>%
 # filter(Year %in% c("2010", "2011", "2012", "2013", "2014", "2015", "2016", "2017"))

#test  <- obs_stan_all_use %>%
 # filter(Year %in% c("2018", "2019", "2020"))

set.seed(3456)
trainIndex <- createDataPartition(obs_stan_all_use$adult_obesity_percent, p = .8, 
                                  list = FALSE, 
                                  times = 1)

train <- obs_stan_all_use[ trainIndex,]
test  <- obs_stan_all_use[-trainIndex,]
```

```{r}
dim(train)#512 rows
dim(test)#192 rows
```


