---
title: "data exploration-food acces"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

packages needed
```{r}
#library(tidyverse)
library(dplyr)
library(tidyr)
library(ggplot2)
library(naniar)
```

read in the data
```{r}
library(readr)
food <- read_csv("Food Access Research Atlas.csv")
#View(food)
```

```{r}
dim(food)
```

only for LA
```{r}
food_la <- food %>%
  filter(food$State == "Louisiana" )
```


```{r}
dim(food_la)
```

number of times a county has a row
```{r}
county_count <- table(food_la$County)
county_count
```

```{r}
summarize(food_la,mean(PovertyRate))
```

avergae poverty rate for each LA county
```{r}
food_county <- group_by(food_la,County)
avg_poverty <- summarize(food_county,avg= mean(PovertyRate,na.rm=TRUE))
avg_poverty
```

average median family income for each LA county
```{r}
avg_income <- summarize(food_county,avg= mean(MedianFamilyIncome,na.rm=TRUE))
avg_income
```


```{r}
avg_pop <- summarize(food_county,avg= mean(POP2010,na.rm=TRUE))
avg_pop
```

LAPOP1_10
LAPOP05_10
LAPOP1_20
LALOWI1_10
LALOWI05_10
LALOWI1_20


```{r}
avg_LAPOP1_10 <- summarize(food_county,avg= mean(LAPOP1_10,na.rm=TRUE))
avg_LAPOP05_10 <- summarize(food_county,avg= mean(LAPOP05_10,na.rm=TRUE))
avg_LAPOP1_20 <- summarize(food_county,avg= mean(LAPOP1_20,na.rm=TRUE))
avg_LALOWI1_10 <- summarize(food_county,avg= mean(LALOWI1_10,na.rm=TRUE))
avg_LALOWI05_10 <- summarize(food_county,avg= mean(LALOWI05_10,na.rm=TRUE))
avg_LALOWI1_20 <- summarize(food_county,avg= mean(LALOWI1_20,na.rm=TRUE))

avg_LAPOP1_10
avg_LAPOP05_10
avg_LAPOP1_20 
avg_LALOWI1_10
avg_LALOWI05_10
avg_LALOWI1_20
```

plotting average income by county
```{r}
ggplot(avg_income, aes(avg, County)) +
 #geom_point(aes(color = LILATracts_Vehicle)) 
  geom_point(aes(color = County))
```


variables flagged for food desert:
LILATracts_1And10
LILATracts_halfAnd10
LILATracts_1And20
LILATracts_Vehicle


relation b/w income and poverty, flagged by vehicular acces
```{r}
ggplot(food_la, aes(MedianFamilyIncome, PovertyRate)) +
 geom_point(aes(color = LILATracts_Vehicle)) 
```

same as above but coloured by county
```{r}
ggplot(food_la, aes(MedianFamilyIncome, PovertyRate)) +
 geom_point(aes(color = County)) 
```


low access and poverty rate

LAPOP1_10
LAPOP05_10
LAPOP1_20
LALOWI1_10
LALOWI05_10
LALOWI1_20
-----------------

lapophalf
lalowihalf
lawhitehalf
lablackhalf
lahisphalf
------------------
lapop1
lalowi1
lawhite1
lablack1
lahisp1
----------------

lapop10
lalowi10
lawhite10
lablack10
lahisp10
----------------------
lapop20: Low access, population at 20 miles, number
lalowi20: Low access, low-income population at 20 miles, number
lawhite20: Low access, White population at 20 miles, number
lablack20: Low access, Black or African American population at 20 miles, number
lahisp20: Low access, Hispanic or Latino population at 20 miles, number



```{r}
#counties from MLR
counties <- c("Assumption", "Cameron", "East Feliciana","Iberville",  "La Salle", "Red River", "West Baton Rouge")

#filter
mlr_counties <- food_la %>%
  filter(County %in% counties)


#adding columns calculating pct of a pop away from supermarket
mlr_counties <- mlr_counties %>%
  #add the overall low acess and low acees low income for urban and rural pops
  mutate(LAPOP1_10_pct = 100*(LAPOP1_10/POP2010), LAPOP05_10_pct = 100*(LAPOP05_10/POP2010), LAPOP1_20_pct = 100*(LAPOP1_20/POP2010), LALOWI1_10_pct = 100*(LALOWI1_10/POP2010), LALOWI05_10_pct = 100*(LALOWI05_10/POP2010), LALOWI1_20_pct = 100*(LALOWI1_20/POP2010),
  
  #individual diantces for low access
  lapophalf_pct = 100*(lapophalf/POP2010), lapop1_pct = 100*(lapop1/POP2010), lapop10_pct = 100*(lapop10/POP2010), lapop20_pct = 100*(lapop20/POP2010),
  
  #individual distances for low access, low income people
  lalowihalf_pct = 100*(lalowihalf/POP2010), lalowi1_pct = 100*(lalowi1/POP2010), lalowi10_pct = 100*(lalowi10/POP2010), lalowi20_pct = 100*(lalowi20/POP2010),
  #distances for racial demographics
  #1/2 mile: lawhitehalf, lablackhalf, lahisphalf
  lawhitehalf_pct = 100*(lawhitehalf/POP2010), lablackhalf_pct = 100*(lablackhalf/POP2010), lahisphalf_pct = 100*(lahisphalf/POP2010),
  
  #1 mile: lawhite1, lablack1, lahisp1
  lawhite1_pct = 100*(lawhite1/POP2010), lablack1_pct = 100*(lablack1/POP2010), lahisp1_pct = 100*(lahisp1/POP2010),
  
  #10 mile: lawhite10, lablack10, lahisp10
  lawhite10_pct = 100*(lawhite10/POP2010), lablack10_pct = 100*(lablack10/POP2010), lahisp10_pct = 100*(lahisp10/POP2010),
  
  #20 mile:
  lawhite20_pct = 100*(lawhite20/POP2010), lablack20_pct = 100*(lablack20/POP2010), lahisp20_pct = 100*(lahisp20/POP2010)) 

```

```{r}
dim(mlr_counties) #31 153 to confirm change
```


```{r}
count <- table(mlr_counties$County)
count
```

avg poverty for each county
```{r}
food_county <- group_by(mlr_counties, County)
avg_poverty <- summarize(food_county,avg= mean(PovertyRate,na.rm=TRUE))
avg_poverty
```

avg income for each county
```{r}
avg_income <- summarize(food_county,avg= mean(MedianFamilyIncome,na.rm=TRUE))
avg_income
```

avg pop for counties
```{r}
avg_pop <- summarize(food_county,avg= mean(POP2010,na.rm=TRUE))
avg_pop
```

LAPOP1_10
LAPOP05_10
LAPOP1_20
LALOWI1_10
LALOWI05_10
LALOWI1_20

calculate averages for rural/urban distnaces

```{r}
avg_LAPOP1_10 <- summarize(food_county,avg= mean(LAPOP1_10_pct,na.rm=TRUE))
avg_LAPOP05_10 <- summarize(food_county,avg= mean(LAPOP05_10_pct,na.rm=TRUE))
avg_LAPOP1_20 <- summarize(food_county,avg= mean(LAPOP1_20_pct,na.rm=TRUE))
avg_LALOWI1_10 <- summarize(food_county,avg= mean(LALOWI1_10_pct,na.rm=TRUE))
avg_LALOWI05_10 <- summarize(food_county,avg= mean(LALOWI05_10_pct,na.rm=TRUE))
avg_LALOWI1_20 <- summarize(food_county,avg= mean(LALOWI1_20_pct,na.rm=TRUE))

avg_LAPOP1_10
avg_LAPOP05_10
avg_LAPOP1_20 
avg_LALOWI1_10
avg_LALOWI05_10
avg_LALOWI1_20
```

#plot avg income
```{r}
ggplot(avg_income, aes(factor(County), avg, fill = factor(County))) +
 #geom_point(aes(color = LILATracts_Vehicle)) 
  #geom_point(aes(color = County))
  geom_bar(stat = "identity") +
  scale_fill_brewer(palette="Dark2")
```

#plot avg poverty
```{r}
ggplot(avg_poverty, aes(factor(County), avg, fill = factor(County))) +
 #geom_point(aes(color = LILATracts_Vehicle)) 
  #geom_point(aes(color = County))
  geom_bar(stat = "identity") +
  scale_fill_brewer(palette="Dark2")
```

avg_LAPOP1_10
avg_LAPOP05_10
avg_LAPOP1_20 
avg_LALOWI1_10
avg_LALOWI05_10
avg_LALOWI1_20

plot low access pop count at different urban/rural distances

```{r}
ggplot(avg_LAPOP1_10, aes(factor(County), avg, fill = factor(County))) +
  geom_bar(stat = "identity")+
        ggtitle("Average low access population count at 1 mile (urban) and 10 miles (rural)")+ 
        #geom_text(aes(label = n_pct))+
        labs(y = "Percentage", x = "")+
        scale_fill_brewer(palette="Dark2")+
        theme_minimal()+
        theme(axis.text.x=element_blank())

ggplot(avg_LAPOP05_10, aes(factor(County), avg, fill = factor(County))) +
  geom_bar(stat = "identity")+
        ggtitle("Average low access population count at 1/2 mile (urban) and 10 miles (rural)")+ 
        #geom_text(aes(label = n_pct))+
        labs(y = "Percentage", x = "")+
        scale_fill_brewer(palette="Dark2")+
        theme_minimal()+
        theme(axis.text.x=element_blank())

ggplot(avg_LAPOP1_20, aes(factor(County), avg, fill = factor(County))) +
  geom_bar(stat = "identity")+
        ggtitle("Average low access population count at 1 mile (urban) and 20 miles (rural)")+ 
        #geom_text(aes(label = n_pct))+
        labs(y = "Percentage", x = "")+
        scale_fill_brewer(palette="Dark2")+
        theme_minimal()+
        theme(axis.text.x=element_blank())

ggplot(avg_LALOWI1_10, aes(factor(County), avg, fill = factor(County))) +
  geom_bar(stat = "identity")+
        ggtitle("Average low access, low income population count at 1 mile (urban) and 10 miles (rural)")+ 
        #geom_text(aes(label = n_pct))+
        labs(y = "Percentage", x = "")+
        scale_fill_brewer(palette="Dark2")+
        theme_minimal()+
        theme(axis.text.x=element_blank())

ggplot(avg_LALOWI05_10, aes(factor(County), avg, fill = factor(County))) +
  geom_bar(stat = "identity")+
        ggtitle("Average low access, low income population count at 1/2 mile (urban) and 10 miles (rural)")+ 
        #geom_text(aes(label = n_pct))+
        labs(y = "Percentage", x = "")+
        scale_fill_brewer(palette="Dark2")+
        theme_minimal()+
        theme(axis.text.x=element_blank())

ggplot(avg_LALOWI1_20, aes(factor(County), avg, fill = factor(County))) +
  geom_bar(stat = "identity")+
        ggtitle("Average low access, low income population count at 1 mile (urban) and 20 miles (rural)")+ 
        #geom_text(aes(label = n_pct))+
        labs(y = "Percentage", x = "")+
        scale_fill_brewer(palette="Dark2")+
        theme_minimal()+
        theme(axis.text.x=element_blank())

```




lapophalf
lapop1
lapop10
lapop20: Low access, population at 20 miles, number

average for just low access at different distances
```{r}
#low access at at 1/2, 1 mile, 10 miles and 20 miles from grocery store
avg_lapophalf <- summarize(food_county,avg= mean(lapophalf_pct,na.rm=TRUE))
avg_lapop1 <- summarize(food_county,avg= mean(lapop1_pct,na.rm=TRUE))
avg_lapop10 <- summarize(food_county,avg= mean(lapop10_pct,na.rm=TRUE))
avg_lapop20 <- summarize(food_county,avg= mean(lapop20_pct,na.rm=TRUE))

avg_lapophalf
avg_lapop1
avg_lapop10 
avg_lapop20
```

use above avergaes for plots
```{r}
ggplot(avg_lapophalf, aes(factor(County), avg, fill = factor(County))) +
  geom_bar(stat = "identity")+
        ggtitle("Average low access population beyond 1/2 mile from supermarket")+ 
        #geom_text(aes(label = n_pct))+
        labs(y = "Percentage", x = "")+
        scale_fill_brewer(palette="Dark2")+
        theme_minimal()+
        theme(axis.text.x=element_blank())

ggplot(avg_lapop1, aes(factor(County), avg, fill = factor(County))) +
  geom_bar(stat = "identity")+
        ggtitle("Average low access population beyond 1 mile from supermarket")+ 
        #geom_text(aes(label = n_pct))+
        labs(y = "Percentage", x = "")+
        scale_fill_brewer(palette="Dark2")+
        theme_minimal()+
        theme(axis.text.x=element_blank())

ggplot(avg_lapop10, aes(factor(County), avg, fill = factor(County))) +
  geom_bar(stat = "identity")+
        ggtitle("Average low access population beyond 10 mile from supermarket")+ 
        #geom_text(aes(label = n_pct))+
        labs(y = "Percentage", x = "")+
        scale_fill_brewer(palette="Dark2")+
        theme_minimal()+
        theme(axis.text.x=element_blank())

ggplot(avg_lapop20, aes(factor(County), avg, fill = factor(County))) +
  geom_bar(stat = "identity")+
        ggtitle("Average low access population beyond 20 mile from supermarket")+ 
        #geom_text(aes(label = n_pct))+
        labs(y = "Percentage", x = "")+
        scale_fill_brewer(palette="Dark2")+
        theme_minimal()+
        theme(axis.text.x=element_blank())
```


lalowihalf
lalowi1
lalowi10
lalowi20: Low access, low-income population at 20 miles, number

```{r}
#low access, low income at 1/2, 1 mile, 10 miles and 20 miles from grocery store
avg_lalowihalf <- summarize(food_county,avg= mean(lalowihalf_pct,na.rm=TRUE))
avg_lalowi1 <- summarize(food_county,avg= mean(lalowi1_pct,na.rm=TRUE))
avg_lalowi10 <- summarize(food_county,avg= mean(lalowi10_pct,na.rm=TRUE))
avg_lalowi20 <- summarize(food_county,avg= mean(lalowi20_pct,na.rm=TRUE))

avg_lalowihalf
avg_lalowi1 
avg_lalowi10
avg_lalowi20
```

avg from above for low acc, low income plots

```{r}
ggplot(avg_lalowihalf, aes(factor(County), avg, fill = factor(County))) +
  geom_bar(stat = "identity")+
        ggtitle("Average low access, low income population beyond 1/2 mile from supermarket")+ 
        #geom_text(aes(label = n_pct))+
        labs(y = "Percentage", x = "")+
        scale_fill_brewer(palette="Dark2")+
        theme_minimal()+
        theme(axis.text.x=element_blank())

ggplot(avg_lalowi1, aes(factor(County), avg, fill = factor(County))) +
  geom_bar(stat = "identity")+
        ggtitle("Average low access, low income population beyond 1 mile from supermarket")+ 
        #geom_text(aes(label = n_pct))+
        labs(y = "Percentage", x = "")+
        scale_fill_brewer(palette="Dark2")+
        theme_minimal()+
        theme(axis.text.x=element_blank())

ggplot(avg_lalowi10, aes(factor(County), avg, fill = factor(County))) +
  geom_bar(stat = "identity")+
        ggtitle("Average low access, low income population beyond 10 mile from supermarket")+ 
        #geom_text(aes(label = n_pct))+
        labs(y = "Percentage", x = "")+
        scale_fill_brewer(palette="Dark2")+
        theme_minimal()+
        theme(axis.text.x=element_blank())

ggplot(avg_lalowi20, aes(factor(County), avg, fill = factor(County))) +
  geom_bar(stat = "identity")+
        ggtitle("Average low access, low income population beyond 20 mile from supermarket")+ 
        #geom_text(aes(label = n_pct))+
        labs(y = "Percentage", x = "")+
        scale_fill_brewer(palette="Dark2")+
        theme_minimal()+
        theme(axis.text.x=element_blank())
```

lawhitehalf
lablackhalf
lahisphalf
```{r}
#races at 1/2 mile from grocery store
avg_lawhitehalf <- summarize(food_county,avg= mean(lawhitehalf_pct,na.rm=TRUE))
avg_lablackhalf <- summarize(food_county,avg= mean(lablackhalf_pct,na.rm=TRUE))
avg_lahisphalf <- summarize(food_county,avg= mean(lahisphalf_pct,na.rm=TRUE))


avg_lawhitehalf
avg_lablackhalf
avg_lahisphalf
```

```{r}
ggplot(avg_lawhitehalf, aes(factor(County), avg, fill = factor(County))) +
  geom_bar(stat = "identity")+
        ggtitle("Average low access white population beyond 1/2 mile from supermarket")+ 
        #geom_text(aes(label = n_pct))+
        labs(y = "Percentage", x = "")+
        scale_fill_brewer(palette="Dark2")+
        theme_minimal()+
        theme(axis.text.x=element_blank())

ggplot(avg_lablackhalf, aes(factor(County), avg, fill = factor(County))) +
  geom_bar(stat = "identity")+
        ggtitle("Average low access black population beyond 1/2 mile from supermarket")+ 
        #geom_text(aes(label = n_pct))+
        labs(y = "Percentage", x = "")+
        scale_fill_brewer(palette="Dark2")+
        theme_minimal()+
        theme(axis.text.x=element_blank())

ggplot(avg_lahisphalf, aes(factor(County), avg, fill = factor(County))) +
  geom_bar(stat = "identity")+
        ggtitle("Average low access hispanic population beyond 1/2 mile from supermarket")+ 
        #geom_text(aes(label = n_pct))+
        labs(y = "Percentage", x = "")+
        scale_fill_brewer(palette="Dark2")+
        theme_minimal()+
        theme(axis.text.x=element_blank())
```


lawhite1
lablack1
lahisp1
```{r}
#races at 1 mile from grocery store
avg_lawhite1 <- summarize(food_county,avg= mean(lawhite1_pct,na.rm=TRUE))
avg_lablack1 <- summarize(food_county,avg= mean(lablack1_pct,na.rm=TRUE))
avg_lahisp1 <- summarize(food_county,avg= mean(lahisp1_pct,na.rm=TRUE))


avg_lawhite1
avg_lablack1
avg_lahisp1
```

```{r}
ggplot(avg_lawhite1, aes(factor(County), avg, fill = factor(County))) +
  geom_bar(stat = "identity")+
        ggtitle("Average low access white population beyond 1 mile from supermarket")+ 
        #geom_text(aes(label = n_pct))+
        labs(y = "Percentage", x = "")+
        scale_fill_brewer(palette="Dark2")+
        theme_minimal()+
        theme(axis.text.x=element_blank())

ggplot(avg_lablack1, aes(factor(County), avg, fill = factor(County))) +
  geom_bar(stat = "identity")+
        ggtitle("Average low access black population beyond 1 mile from supermarket")+ 
        #geom_text(aes(label = n_pct))+
        labs(y = "Percentage", x = "")+
        scale_fill_brewer(palette="Dark2")+
        theme_minimal()+
        theme(axis.text.x=element_blank())

ggplot(avg_lahisp1, aes(factor(County), avg, fill = factor(County))) +
  geom_bar(stat = "identity")+
        ggtitle("Average low access hispanic population beyond 1 mile from supermarket")+ 
        #geom_text(aes(label = n_pct))+
        labs(y = "Percentage", x = "")+
        scale_fill_brewer(palette="Dark2")+
        theme_minimal()+
        theme(axis.text.x=element_blank())
```

lawhite10
lablack10
lahisp10

```{r}
#races at 10 mile from grocery store
avg_lawhite10 <- summarize(food_county,avg= mean(lawhite10_pct,na.rm=TRUE))
avg_lablack10 <- summarize(food_county,avg= mean(lablack10_pct,na.rm=TRUE))
avg_lahisp10 <- summarize(food_county,avg= mean(lahisp10_pct,na.rm=TRUE))


avg_lawhite10
avg_lablack10
avg_lahisp10
```


```{r}
ggplot(avg_lawhite10, aes(factor(County), avg, fill = factor(County))) +
  geom_bar(stat = "identity")+
        ggtitle("Average low access white population beyond 10 mile from supermarket")+ 
        #geom_text(aes(label = n_pct))+
        labs(y = "Percentage", x = "")+
        scale_fill_brewer(palette="Dark2")+
        theme_minimal()+
        theme(axis.text.x=element_blank())

ggplot(avg_lablack10, aes(factor(County), avg, fill = factor(County))) +
  geom_bar(stat = "identity")+
        ggtitle("Average low access black population beyond 10 mile from supermarket")+ 
        #geom_text(aes(label = n_pct))+
        labs(y = "Percentage", x = "")+
        scale_fill_brewer(palette="Dark2")+
        theme_minimal()+
        theme(axis.text.x=element_blank())

ggplot(avg_lahisp10, aes(factor(County), avg, fill = factor(County))) +
  geom_bar(stat = "identity")+
        ggtitle("Average low access hispanic population beyond 10 mile from supermarket")+ 
        #geom_text(aes(label = n_pct))+
        labs(y = "Percentage", x = "")+
        scale_fill_brewer(palette="Dark2")+
        theme_minimal()+
        theme(axis.text.x=element_blank())
```

lawhite20: Low access, White population at 20 miles, number
lablack20: Low access, Black or African American population at 20 miles, number
lahisp20: Low access, Hispanic or Latino population at 20 miles, number
```{r}
#races at 20 mile from grocery store
avg_lawhite20 <- summarize(food_county,avg= mean(lawhite20_pct,na.rm=TRUE))
avg_lablack20 <- summarize(food_county,avg= mean(lablack20_pct,na.rm=TRUE))
avg_lahisp20 <- summarize(food_county,avg= mean(lahisp20_pct,na.rm=TRUE))


avg_lawhite20
avg_lablack20
avg_lahisp20
```

```{r}
ggplot(avg_lawhite20, aes(factor(County), avg, fill = factor(County))) +
  geom_bar(stat = "identity")+
        ggtitle("Average low access white population beyond 20 mile from supermarket")+ 
        #geom_text(aes(label = n_pct))+
        labs(y = "Percentage", x = "")+
        scale_fill_brewer(palette="Dark2")+
        theme_minimal()+
        theme(axis.text.x=element_blank())

ggplot(avg_lablack20, aes(factor(County), avg, fill = factor(County))) +
  geom_bar(stat = "identity")+
        ggtitle("Average low access black population beyond 20 mile from supermarket")+ 
        #geom_text(aes(label = n_pct))+
        labs(y = "Percentage", x = "")+
        scale_fill_brewer(palette="Dark2")+
        theme_minimal()+
        theme(axis.text.x=element_blank())

ggplot(avg_lahisp20, aes(factor(County), avg, fill = factor(County))) +
  geom_bar(stat = "identity")+
        ggtitle("Average low access hispanic population beyond 20 mile from supermarket")+ 
        #geom_text(aes(label = n_pct))+
        labs(y = "Percentage", x = "")+
        scale_fill_brewer(palette="Dark2")+
        theme_minimal()+
        theme(axis.text.x=element_blank())
```



